On the uniqueness and stability of dictionaries for sparse representation of noisy signals

### 2017-04-06 manuscripts/PNAS/chazthm_pnas.tex
sml edts

\section{Results}

@@

\begin{lemma}\label{NonEmptyLemma} 
Fix $\ell, m$, and $\bar m$. Suppose $E \subseteq 2^{[m]}$ is $\ell$-regular with the SIP.  If $\pi: E \to 2^{[\bar m]}$ has $\sum_{S \in E} |S| = \sum_{S \in E} |\pi(S)|$ and:
\begin{align}\label{cond}
|\bigcap_{S \in F} \pi(S)| \leq |\bigcap_{S \in F} S |,\ \ \   \text{for } F \subseteq E,
\end{align}
%
then $\bar m \geq m$ and the association $i \mapsto \cap \pi(F(i))$ defines an injective map from $J$ to $[\bar m]$ for some $J \subseteq [m]$ of size $\bar m - \ell(\bar m - m)$. In particular, if $\bar m = m$ then the map $\pi$ is induced by a permutation.
%$\pi(E)$ is $\ell$-regular and satisfies the singleton intersection property. In particular, $|\cap_{S \in F} S| = 1$ if and only if $|\cap_{S \in F} \pi(S)| = 1$. 
[*** TODO: Chris to Finish ***]
\end{lemma}
\begin{proof}
  We first show that $\bar m \geq m$. Fix $T = \{(j, S): j \in \pi(S), S \in E\}$. Then $|T| = \sum_{S \in E} |\pi(S)| = \sum_{S \in E} |S| = \sum_{i \in [m]} \deg(i) = \ell m$ by regularity of $E$. If $\bar m < m$ then by pigeonholing the $m \ell \geq \ell (\bar m + 1) \geq [(\ell + 1) - 1] \bar m + 1$ elements of $T$ with respect to the set $[\bar m]$ of possible first indices, we see that there must exist at least $\ell + 1$ elements of $T$ having the same first index. But by \eqref{cond} there can be no more than $\ell$ elements of $T$ with a given first index, since $E$ is $\ell$-regular. Hence $\bar m \geq m$. 
Again, pigeonholing the elements of $T$ with respect to the set $[\bar m]$ of possible first indices, we see that for each $j$ there must be exactly $\ell$ elements of $T$ with $j$ as a first index; hence, $\pi(E)$ is $\ell$-regular. 
Fix $j$ and let $F(j) = \{S \in E: j \in \pi(S)\}$ be the preimage of the star in $\pi(E)$ centered at $j$, which is of size $\ell$ by the above arguments. It follows from \eqref{cond} that $\cap_{S \in F(j)} S$ is nonempty. In fact, we must have $|\cap _{S \in F(j)} S| = 1$, since $E$ is $\ell$-regular and satisfies the singleton intersection property. It follows by \eqref{cond} that $\cap_{S \in F(j)} \pi(S) = \{j\}$. Thus $\pi(E)$ satisfies the singleton intersection property as well, and the preimage of every one of the $m$ stars in $\pi(E)$ is a star in $E$. It remains to show that every star in $E$ maps through $\pi$ to some star in $\pi(E)$. This follows by pigeonholing the $m$ stars of $\pi(E)$ with respect to their $m$ possible preimages if no two stars in $\pi(E)$ share the same preimage. This indeed must be the case, since $F(i) = F(j)$ implies $\{i\} = \cap_{S \in F(i)} \pi(S) = \cap_{S \in F(j)} \pi(S) = \{j\}$.
\end{proof}

@@

\section{Proof (sketches) of Thm.~\ref{robustPolythm} and Cor.~\ref{ProbabilisticCor}}\label{AppendixB} %Cors.~\ref{DeterministicUniquenessCorollary} \& \ref{ProbabilisticCor}}\label{AppendixB}

@@

\begin{proof}[Proof (sketch) of Thm.~\ref{robustPolythm}]
% We sketch the argument, leaving the details to the reader. 
Let $M$ be the matrix with columns $\mathbf{A}\mathbf{x}_i$, $i \in [N]$.  Consider the polynomial \cite[Sec.~IV]{Hillar15} in the entries of $\mathbf{A}$ and $\mathbf{x}_i$:
\begin{align*}
g(\mathbf{A}, \{\mathbf{x}_i\}_{i=1}^N) = \prod_{S \in {[n] \choose k}} \sum_{S' \in {[N] \choose k}} (\det M_{S',S})^2,
\end{align*}
with notation as in Sec.~\ref{Results}.  
It can be checked that when $g$ is nonzero for a substitution of real numbers for the indeterminates, all of the genericity requirements on $\mathbf{A}$ and $\mathbf{x}_i$ in our proofs of stability in Thm.~\ref{DeterministicUniquenessTheorem} are satisfied (in particular, the spark condition on $\mathbf{A}$). % The statement of the theorem now follows directly.
\end{proof}

\begin{proof}[Proof of Cor.~\ref{ProbabilisticCor}]
First, note that if a set of measure spaces $\{(X_{\ell}, \Sigma_{\ell}, \nu_{\ell})\}_{\ell=1}^p$ is such that $\nu_{\ell}$ is absolutely continuous with respect to $\mu$ for all $\ell = 1, \ldots, p$, where $\mu$ is the standard Borel measure on $\mathbb{R}$, then the product measure $\prod_{\ell=1}^p \nu_{\ell}$ is absolutely continuous with respect to the standard Borel product measure on $\mathbb{R}^p$ (e.g.,  \cite{folland2013real}). By Thm.~\ref{robustPolythm}, there is a polynomial that is nonzero whenever $Y$ has a stable $k$-sparse representation in $\mathbb R^m$; in particular, this property (stability) holds with probability one.
\end{proof}

>>>>

\section{Results}\label{results}

@@

\begin{lemma}\label{NonEmptyLemma} 
Fix $\ell, m$, and $\bar m$. Suppose $E \subseteq 2^{[m]}$ is $\ell$-regular with the SIP.  If $\pi: E \to 2^{[\bar m]}$ has $\sum_{S \in E} |S| = \sum_{S \in E} |\pi(S)|$ and:
\begin{align}\label{cond}
|\bigcap_{S \in F} \pi(S)| \leq |\bigcap_{S \in F} S |,\ \ \   \text{for } F \subseteq E,
\end{align}
%
then $\bar m \geq m$ and the association $i \mapsto \cap \pi(F(i))$ defines an injective map from $J$ to $[\bar m]$ for some $J \subseteq [m]$ of size $\bar m - \ell(\bar m - m)$. In particular, if $\bar m = m$ then the map $\pi$ is induced by a permutation.
%$\pi(E)$ is $\ell$-regular and satisfies the singleton intersection property. In particular, $|\cap_{S \in F} S| = 1$ if and only if $|\cap_{S \in F} \pi(S)| = 1$. 
[*** TODO: Insert the final proof here ***]
\end{lemma}
\begin{proof}
  We first show that $\bar m \geq m$. Fix $T = \{(j, S): j \in \pi(S), S \in E\}$. Then $|T| = \sum_{S \in E} |\pi(S)| = \sum_{S \in E} |S| = \sum_{i \in [m]} \deg(i) = \ell m$ by regularity of $E$. If $\bar m < m$ then by pigeonholing the $m \ell \geq \ell (\bar m + 1) \geq [(\ell + 1) - 1] \bar m + 1$ elements of $T$ with respect to the set $[\bar m]$ of possible first indices, we see that there must exist at least $\ell + 1$ elements of $T$ having the same first index. But by \eqref{cond} there can be no more than $\ell$ elements of $T$ with a given first index, since $E$ is $\ell$-regular. Hence $\bar m \geq m$. 
Again, pigeonholing the elements of $T$ with respect to the set $[\bar m]$ of possible first indices, we see that for each $j$ there must be exactly $\ell$ elements of $T$ with $j$ as a first index; hence, $\pi(E)$ is $\ell$-regular. 
Fix $j$ and let $F(j) = \{S \in E: j \in \pi(S)\}$ be the preimage of the star in $\pi(E)$ centered at $j$, which is of size $\ell$ by the above arguments. It follows from \eqref{cond} that $\cap_{S \in F(j)} S$ is nonempty. In fact, we must have $|\cap _{S \in F(j)} S| = 1$, since $E$ is $\ell$-regular and satisfies the singleton intersection property. It follows by \eqref{cond} that $\cap_{S \in F(j)} \pi(S) = \{j\}$. Thus $\pi(E)$ satisfies the singleton intersection property as well, and the preimage of every one of the $m$ stars in $\pi(E)$ is a star in $E$. It remains to show that every star in $E$ maps through $\pi$ to some star in $\pi(E)$. This follows by pigeonholing the $m$ stars of $\pi(E)$ with respect to their $m$ possible preimages if no two stars in $\pi(E)$ share the same preimage. This indeed must be the case, since $F(i) = F(j)$ implies $\{i\} = \cap_{S \in F(i)} \pi(S) = \cap_{S \in F(j)} \pi(S) = \{j\}$.
\end{proof}

@@

\section{Proofs of Thm.~\ref{robustPolythm} and Cor.~\ref{ProbabilisticCor}}\label{AppendixB} %Cors.~\ref{DeterministicUniquenessCorollary} \& \ref{ProbabilisticCor}}\label{AppendixB}

@@

\begin{proof}[Proof (sketch) of Thm.~\ref{robustPolythm}]
% We sketch the argument, leaving the details to the reader.  %\cite[Sec.~IV]{Hillar15}
Let $M$ be the matrix with columns $\mathbf{A}\mathbf{x}_i$, $i \in [N]$.  Consider the following polynomial in the entries of $\mathbf{A}$, $\mathbf{x}_i$:
\begin{align*}
g(\mathbf{A}, \{\mathbf{x}_i\}_{i=1}^N) = \prod_{S \in {[n] \choose k}} \sum_{S' \in {[N] \choose k}} (\det M_{S',S})^2,
\end{align*}
with notation as in Sec.~\ref{results}.  It can be checked that when $g$ is nonzero for a substitution of real numbers for the indeterminates, all of the genericity requirements on $\mathbf{A}$ and $\mathbf{x}_i$ in our proofs of stability in Thm.~\ref{DeterministicUniquenessTheorem} are satisfied (in particular, the spark condition on $\mathbf{A}$). % The statement of the theorem now follows directly.
\end{proof}

\begin{proof}[Proof (sketch) of Cor.~\ref{ProbabilisticCor}]
First, note that if a set of measure spaces $\{(X_{\ell}, \Sigma_{\ell}, \nu_{\ell})\}_{\ell=1}^p$ has $\nu_{\ell}$ is absolutely continuous with respect to $\mu$ for all $\ell \in [p]$, where $\mu$ is the standard Borel measure on $\mathbb{R}$, then the product measure $\prod_{\ell=1}^p \nu_{\ell}$ is absolutely continuous with respect to the standard Borel product measure on $\mathbb{R}^p$ (e.g.,  \cite{folland2013real}). By Thm.~\ref{robustPolythm}, there is a polynomial that is nonzero whenever $Y$ has a stable $k$-sparse representation in $\mathbb R^m$; in particular, this property (stability) holds with probability one.
\end{proof}
